---
title: "cLFM Tutorial"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Install missing packages
list.of.packages <- c("common", "dplyr", "fANCOVA", "fda", "fdapace",
                      "latex2exp", "mgcv", "pracma", "psych", "tidyverse")
new.packages <- list.of.packages[!(list.of.packages %in% 
                                     installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(common)
library(dplyr)
library(fANCOVA)
library(fda)
library(fdapace)
library(latex2exp)
library(mgcv)
library(pracma)
library(psych)
library(tidyverse)
rm(new.packages, list.of.packages)
```


# Introduction

## Description

A step-by-step implementation of cLFM and associated procedures including data 
generating process of the first simulation scenarios and evaluation of model 
parameter estimation. Details of simulation design, EM estimation of model 
parameters and trimming-refinement process are described in Section 3 and 
Section 4 of 'Contrastive Latent Functional Model'.

## Main functions implemented

1. data_gen: Function that generates a pair of data sets composed of preset shared and 
unique variation patterns.
2. cLFM: Function that constructs and fits the proposed contrastive latent functional model.

## Required files

1. Simulation_dgp.R
2. cLFM_Functions_1.R
3. cLFM_Functions_2.R


# Implementation

## 1. Load the required files

```{r, warning=FALSE}
source("Simulation_dgp.R")
source("cLFM_Functions_1.R")
source("cLFM_Functions_2.R")
```


## 2. cLFM implementation

```{r}
# Containers to save the estimates of overall mean functions, error variances, latent components and score variances.
mu_x_save = list()
mu_y_save = list()
s2_save = list()

Psi_save = list()
Phi_save = list()
Gamma_save = list()

Omega_x_save = list()
Omega_y_save = list()
Theta_save = list()
Lambda_save = list()
```

```{R}
for (ite in 1:10) {
  print(paste0(ite, "-th iteration starts here ------------------------------"))
  # Generate the simulation data pairs
  data_grids = data_gen(n_x = 50, n_y = 50, s2_x = 5, s2_y = 5)
  X = data_grids[[1]]
  Y = data_grids[[2]]
  
  # Fitting the proposed cLFM through the EM estimation and Trimming-Refinement process
  results = cLFM(X = X, Y = Y, alpha_L = .85, alpha_KR = .85,
                 alpha_refine = .85, verbose = TRUE)
  
  # Save the estimates
  mu_x_save = append(mu_x_save, list(results[[1]]))
  mu_y_save = append(mu_y_save, list(results[[2]]))
  s2_save = append(s2_save, list(c(results[[3]], results[[4]])))
  
  Psi_save = append(Psi_save, list(results[[5]]))
  Phi_save = append(Phi_save, list(results[[6]]))
  Gamma_save = append(Gamma_save, list(results[[7]]))
  
  Omega_x_save = append(Omega_x_save, list(results[[8]]))
  Omega_y_save = append(Omega_y_save, list(results[[9]]))
  Theta_save = append(Theta_save, list(results[[10]]))
  Lambda_save = append(Lambda_save, list(results[[11]]))
}
```


## 3. Evaluation

True model parameters

```{r, warning=FALSE}
N = 30
tobs = seq(0, 1, length.out = N)
# True mean function mu^x(t)
mu_x = matrix(sapply(tobs, function(x){return(4.5 * x + 1.5 * exp(x ^ 2))}), 
              nrow = N, ncol = 1) 
# True mean function mu^y(t)
mu_y = matrix(sapply(tobs, function(x){return(-10 * x + 5 * 
                                                exp(-50 * (x - 0.7) ^ 2) + 5)}), 
              nrow = N, ncol = 1)
# True error variances
s2_x = 5
s2_y = 5
# True latent components and score variances
Psi_true = eval.basis(tobs, fda::create.fourier.basis(nbasis = 3,
                                                      dropind = c(1,2)))
Phi_true = eval.basis((tobs), create.fourier.basis(nbasis = 3,
                                                   dropind = c(1,3)))
Gamma_true = eval.basis((tobs), create.fourier.basis(nbasis = 5,
                                                     dropind = c(1,2,3,5)))
Omega_x = matrix(14)
Omega_y = matrix(11.5)
Theta = matrix(6)
Lambda = matrix(8.5)
```

MSDE of overall mean functions $\mu^x(t)$, $\mu^y(t)$:

```{r}
MSDE_mu = matrix(data = NA, nrow = length(mu_x_save), ncol = 2)
for (i in 1:length(mu_x_save)) {
  MSDE_mu[i,1] = trapz(tobs, (mu_x - mu_x_save[[i]])^2) / trapz(tobs, (mu_x)^2)
  MSDE_mu[i,2] = trapz(tobs, (mu_y - mu_y_save[[i]])^2) / trapz(tobs, (mu_y)^2)
}
print(paste0("The mean MSDE for the overall mean functions {", 
             symbol("mu") %p% supsc("x"), "(t), ", 
             symbol("mu") %p% supsc("y"), "(t)} are:"))
print(round(colMeans(MSDE_mu, na.rm = TRUE), digits = 3))
```

MSE of error variances $\sigma^2_X$, $\sigma^2_Y$:

```{r}
MSE_s2 = matrix(data = NA, nrow = length(s2_save), ncol = 2)
for (i in 1:length(s2_save)) {
  MSE_s2[i,1] = (s2_x - s2_save[[i]][1])^2 / s2_x^2
  MSE_s2[i,2] = (s2_y - s2_save[[i]][2])^2 / s2_y^2
}
print(paste0("The mean MSE for the error variances {", 
             symbol("sigma") %p% supsc("2"), "X, ", 
             symbol("sigma") %p% supsc("2"), "Y} are:"))
print(round(colMeans(MSE_s2, na.rm = TRUE), digits = 3))
```



Calculate MSDE for latent component estimates and MSE for score variance estimates, and plot true and 
estimated trajectories with (5th, 50th, 95th) MSDE values.

Psi:

```{r}
# When more than 1 shared components are used in data generating process or estimated by cLFM, we calculate 
# the MSDE between each pair of the true and estimated components to determine their correspondence.
L_true = 1 # only one shared component is used in simulation scenario 1 to generate data pairs.
MSDE_Psi = matrix(data = NA, nrow = length(Psi_save), ncol = L_true)
Psi_1_est = matrix(data = NA, nrow = length(Psi_save), ncol = N)
Psi_2_est = matrix(data = NA, nrow = length(Psi_save), ncol = N)
Omega_x_est = matrix(data = NA, nrow = length(Omega_x_save), ncol = L_true)
Omega_y_est = matrix(data = NA, nrow = length(Omega_y_save), ncol = L_true)
for (i in 1:length(Psi_save)) {
  if(!is.null(Psi_save[[i]])){ # skipped if no shared components are estimated.
    Psi_est = Psi_save[[i]]
    if(ncol(Psi_est) == 1 & L_true == 1){ # L_true = L_est = 1, the correspondence is fixed.
      # Considering possible flipping forms between the estimated and true components, we take the
      # minimal MSDE value from following difference integration.
      MSDE_Psi[i,1] = min(
        trapz(tobs, (Psi_true - Psi_est)^2)/trapz(tobs, Psi_true^2), 
        trapz(tobs, (Psi_true + Psi_est)^2)/trapz(tobs, Psi_true^2))
      Psi_1_est[i,] = Psi_est
      Omega_x_est[i,1] = Omega_x_save[[i]]
      Omega_y_est[i,1] = Omega_y_save[[i]]
    }else if(L_true == 1){ # L_est > L_true = 1
      # When number of shared component estimates returned are larger than the number of true shared component, 
      # the one having the lowest MSDE with the true component is referred to as its correspondence.
      for (j in 1:ncol(Psi_est)) {
        MSDE = min(trapz(tobs, (Psi_true - Psi_est[,j])^2) / 
                     trapz(tobs, Psi_true^2), 
                 trapz(tobs, (Psi_true + Psi_est[,j])^2) / 
                   trapz(tobs, Psi_true^2))
        if(j > 1){
          if(MSDE < MSDE_Psi[i, 1]){
            MSDE_Psi[i, 1] = MSDE
            Psi_1_est[i,] = Psi_est[,j]
            Omega_x_est[i,1] = Omega_x_save[[i]][j,j]
            Omega_y_est[i,1] = Omega_y_save[[i]][j,j]
          }
        }else{
          MSDE_Psi[i, 1] = MSDE
          Psi_1_est[i,] = Psi_est[,j]
          Omega_x_est[i,1] = Omega_x_save[[i]][j,j]
          Omega_y_est[i,1] = Omega_y_save[[i]][j,j]
        }
      }
    }else if(ncol(Psi_est) == 1){ # L_true > L_est = 1
        MSDE_1 = min(trapz(tobs, (Psi_true[,1] - Psi_est)^2) / 
                       trapz(tobs, Psi_true[,1]^2), 
                 trapz(tobs, (Psi_true[,1] + Psi_est)^2) / 
                   trapz(tobs, Psi_true[,1]^2))
        MSDE_2 = min(trapz(tobs, (Psi_true[,2] - Psi_est)^2) / 
                       trapz(tobs, Psi_true[,2]^2), 
                 trapz(tobs, (Psi_true[,2] + Psi_est)^2) / 
                   trapz(tobs, Psi_true[,2]^2))
        if(MSDE_1 < MSDE_2){
          MSDE_Psi[i, 1] = MSDE_1
          Psi_1_est[i,] = Psi_est
          Omega_x_est[i,1] = Omega_x_save[[i]]
          Omega_y_est[i,1] = Omega_y_save[[i]]
          MSDE_Psi[i, 2] = NA
          Omega_x_est[i,2] = NA
          Omega_y_est[i,2] = NA
        }else{
          MSDE_Psi[i, 2] = MSDE_2
          Psi_2_est[i,] = Psi_est
          Omega_x_est[i,2] = Omega_x_save[[i]]
          Omega_y_est[i,2] = Omega_y_save[[i]]
          MSDE_Psi[i, 1] = NA
          Omega_x_est[i,1] = NA
          Omega_y_est[i,1] = NA
        }
    }else{ # L_true > 1, L_est > 1.
      ind_psi_1 = 0 # used to save the index of estimated component corresponding to psi_1(t)
      MSDE_array = c()
      for (j in 1:ncol(Psi_est)) {
        MSDE_array = append(MSDE_array, min(
          trapz(tobs, (Psi_true[,1] - Psi_est[,j])^2) / 
            trapz(tobs, Psi_true[,1]^2),
          trapz(tobs, (Psi_true[,1] + Psi_est[,j])^2) / 
            trapz(tobs, Psi_true[,1]^2)))
        ind_psi_1 = which.min(MSDE_array)
      }
      
      ind_psi_2 = 0 # used to save the index of estimated component corresponding to psi_2(t)
      MSDE_array = c()
      for (j in 1:ncol(Psi_est)) {
        MSDE_array = append(MSDE_array, min(
          trapz(tobs, (Psi_true[,2] - Psi_est[,j])^2) / 
            trapz(tobs, Psi_true[,2]^2), 
          trapz(tobs, (Psi_true[,2] + Psi_est[,j])^2) / 
            trapz(tobs, Psi_true[,2]^2)))
        ind_psi_2 = which.min(MSDE_array)
      }
      
      if(ind_psi_1 != ind_psi_2){
        MSDE_Psi[i, 1] = min(
          trapz(tobs, (Psi_true[,1] - Psi_est[,ind_psi_1])^2) / 
            trapz(tobs, Psi_true[,1]^2), 
          trapz(tobs, (Psi_true[,1] + Psi_est[,ind_psi_1])^2) / 
            trapz(tobs, Psi_true[,1]^2))
        Psi_1_est[i,] = Psi_est[,ind_psi_1]
        Omega_x_est[i,1] = Omega_x_save[[i]][ind_psi_1,ind_psi_1]
        Omega_y_est[i,1] = Omega_y_save[[i]][ind_psi_1,ind_psi_1]
        
        MSDE_Psi[i, 2] = min(
          trapz(tobs, (Psi_true[,2] - Psi_est[,ind_psi_2])^2) / 
            trapz(tobs, Psi_true[,2]^2), 
          trapz(tobs, (Psi_true[,2] + Psi_est[,ind_psi_2])^2) / 
            trapz(tobs, Psi_true[,2]^2))
        Psi_2_est[i,] = Psi_est[,ind_psi_2]
        Omega_x_est[i,2] = Omega_x_save[[i]][ind_psi_2,ind_psi_2]
        Omega_y_est[i,2] = Omega_y_save[[i]][ind_psi_2,ind_psi_2]
      }else{
        print(paste0("In ", i, "th Monte Carlo run, the two true shared ", 
                     "components are corresponding to the same estimate."))
      }
    }
  }
}

print(paste0("The mean MSDE for the shared components {", symbol("psi"), 
             "(t)} ", "are:", round(mean(MSDE_Psi[,1], na.rm = TRUE), 
                                    digits = 3)))
if(L_true > 1){
  print(round(mean(MSDE_Psi[,2], na.rm = TRUE), digits = 3))
}


# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = L_true, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
quantile_mat[1,] = sort(na.omit(MSDE_Psi)[,1])[ceiling(nrow(na.omit(MSDE_Psi)) * 
                                                       c(.05, .50, .95))] 
if(L_true > 1){
  quantile_mat[2,] = sort(na.omit(MSDE_Psi)[,2])[
    ceiling(nrow(na.omit(MSDE_Psi)) * c(.05, .50, .95))]
}

ind_mat = matrix(nrow = L_true, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for (j in 1:3) {
  ind_mat[1,j] = which.min(abs(MSDE_Psi[,1] - quantile_mat[1,j]))
  if(L_true > 1){
    ind_mat[2,j] = which.min(abs(MSDE_Psi[,2] - quantile_mat[2,j]))
  }
}
psi_1_5th = sign(Psi_1_est[ind_mat[1,1],4]) * Psi_1_est[ind_mat[1,1],]
psi_1_50th = sign(Psi_1_est[ind_mat[1,2],4]) * Psi_1_est[ind_mat[1,2],]
psi_1_95th = sign(Psi_1_est[ind_mat[1,3],4]) * Psi_1_est[ind_mat[1,3],]


## plot for psi_1(t)
matplot(tobs, cbind(Psi_true[,1], psi_1_5th, psi_1_50th, psi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), xaxs = 'i')
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(psi)[1]*(t)), side=2, line=1.9, cex=1.9)
legend("bottomright", c("true", "5th", "50th", "95th"), lty = c(1,2,3,4), 
       col=c("black", "red", "black", "blue"), lwd = 2, cex=1.2)


## plot for psi_2(t)
if(L_true > 1){
  psi_2_5th = sign(Psi_2_est[ind_mat[1,1],4]) * Psi_2_est[ind_mat[1,1],]
  psi_2_50th = sign(Psi_2_est[ind_mat[1,2],4]) * Psi_2_est[ind_mat[1,2],]
  psi_2_95th = sign(Psi_2_est[ind_mat[1,3],4]) * Psi_2_est[ind_mat[1,3],]
  matplot(tobs, cbind(Psi_true[,2], psi_2_5th, psi_2_50th, psi_2_95th), 
          type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
          cex.axis = 1.5, col=c("black", "red","black", "blue"), 
          lty = c(1,2,3,4), mgp = c(2.5,1,0), xaxs = 'i')
  title(main = list("", cex = 2.5), adj = 0, line = 1.5)
  mtext(expression(hat(psi)[2]*(t)), side=2, line=1.9, cex=1.9)
}
```



Phi

```{r}
K_true = 1 # only one unique component is used in simulation scenario 1 to generate X.

# MSDE

MSDE_Phi = matrix(data = NA, nrow = length(Phi_save), ncol = K_true)
Phi_1_est = matrix(data = NA, nrow = length(Phi_save), ncol = N)
Phi_2_est = matrix(data = NA, nrow = length(Phi_save), ncol = N)
Theta_est = matrix(data = NA, nrow = length(Theta_save), ncol = K_true)
for (i in 1:length(Phi_save)) {
  if(!is.null(Phi_save[[i]])){ # skipped if no unique components are estimated.
    Phi_est = Phi_save[[i]]
    if(ncol(Phi_est) == 1 & K_true == 1){ # K_true = K_est = 1, the correspondence is fixed.
      # Considering possible flipping forms between the estimated and true components, we take 
      # the minimal MSDE value from following difference integration.
      MSDE_Phi[i,1] = min(
        trapz(tobs, (Phi_true - Phi_est)^2) / trapz(tobs, Phi_true^2), 
        trapz(tobs, (Phi_true + Phi_est)^2) / trapz(tobs, Phi_true^2))
      Phi_1_est[i,] = Phi_est
      Theta_est[i,1] = Theta_save[[i]]
    }else if(K_true == 1){ # K_est > K_true = 1
      # When the number of unique component estimates returned are greater than the true number, 
      # the one having the lowest MSDE with the true component is referred to as its correspondence.
      for (j in 1:ncol(Phi_est)) {
        MSDE = min(trapz(tobs, (Phi_true - Phi_est[,j])^2) / 
                     trapz(tobs, Phi_true^2), 
                 trapz(tobs, (Phi_true + Phi_est[,j])^2) / 
                   trapz(tobs, Phi_true^2))
        if(j > 1){
          if(MSDE < MSDE_Phi[i, 1]){
            MSDE_Phi[i, 1] = MSDE
            Phi_1_est[i,] = Phi_est[,j]
            Theta_est[i,1] = Theta_save[[i]][j,j]
          }
        }else{
          MSDE_Phi[i, 1] = MSDE
          Phi_1_est[i,] = Phi_est[,j]
          Theta_est[i,1] = Theta_save[[i]][j,j]
        }
      }
    }else if(ncol(Phi_est) == 1){ # K_true > K_est = 1
        MSDE_1 = min(trapz(tobs, (Phi_true[,1] - Phi_est)^2) / 
                       trapz(tobs, Phi_true[,1]^2), 
                 trapz(tobs, (Phi_true[,1] + Phi_est)^2) / 
                   trapz(tobs, Phi_true[,1]^2))
        MSDE_2 = min(trapz(tobs, (Phi_true[,2] - Phi_est)^2) / 
                       trapz(tobs, Phi_true[,2]^2), 
                 trapz(tobs, (Phi_true[,2] + Phi_est)^2) / 
                   trapz(tobs, Phi_true[,2]^2))
        if(MSDE_1 < MSDE_2){
          MSDE_Phi[i,1] = MSDE_1
          Phi_1_est[i,] = Phi_est
          Theta_est[i,1] = Theta_save[[i]]
          MSDE_Phi[i,2] = NA
          Theta_est[i,2] = NA
        }else{
          MSDE_Phi[i,2] = MSDE_2
          Phi_2_est[i,] = Phi_est
          Theta_est[i,2] = Theta_save[[i]]
          MSDE_Phi[i,1] = NA
          Theta_est[i,1] = NA
        }
    }else{ # K_true > 1, K_est > 1
      ind_phi_1 = 0 # used to record the index of estimated component corresponding to phi_1(t)
      MSDE_array = c()
      for (j in 1:ncol(Phi_est)) {
        MSDE_array = append(MSDE_array, min(
          trapz(tobs, (Phi_true[,1] - Phi_est[,j])^2) / 
            trapz(tobs, Phi_true[,1]^2),
          trapz(tobs, (Phi_true[,1] + Phi_est[,j])^2) / 
            trapz(tobs, Phi_true[,1]^2)))
        ind_phi_1 = which.min(MSDE_array)
      }
      
      ind_phi_2 = 0 # used to record the index of estimated component corresponding to phi_2(t)
      MSDE_array = c()
      for (j in 1:ncol(Phi_est)) {
        MSDE_array = append(MSDE_array, min(
          trapz(tobs, (Phi_true[,2] - Phi_est[,j])^2) / 
            trapz(tobs, Phi_true[,2]^2), 
          trapz(tobs, (Phi_true[,2] + Phi_est[,j])^2) / 
            trapz(tobs, Phi_true[,2]^2)))
        ind_phi_2 = which.min(MSDE_array)
      }
      
      if(ind_phi_1 != ind_phi_2){
        MSDE_Phi[i,1] = min(trapz(tobs, 
                                   (Phi_true[,1] - Phi_est[,ind_phi_1])^2) / 
                               trapz(tobs, Phi_true[,1]^2), 
                 trapz(tobs, (Phi_true[,1] + Phi_est[,ind_phi_1])^2) / 
                   trapz(tobs, Phi_true[,1]^2))
        Phi_1_est[i,] = Phi_est[,ind_phi_1]
        Theta_est[i,1] = Theta_save[[i]][ind_phi_1,ind_phi_1]
        
        MSDE_Phi[i,2] = min(trapz(tobs, 
                                   (Phi_true[,2] - Phi_est[,ind_phi_2])^2) / 
                               trapz(tobs, Phi_true[,2]^2), 
                 trapz(tobs, (Phi_true[,2] + Phi_est[,ind_phi_2])^2) / 
                   trapz(tobs, Phi_true[,2]^2))
        Phi_2_est[i,] = Phi_est[,ind_phi_2]
        Theta_est[i,2] = Theta_save[[i]][ind_phi_2,ind_phi_2]
      }else{
        print(paste0("In ", i, "th Monte Carlo run, the two true unique ", 
                     "components are corresponding to the same estimate."))
      }
    }
  }
}


print(paste0("The mean MSDE for the unique components {", symbol("phi"), 
             "(t)} ", "are:", round(mean(MSDE_Phi[,1], na.rm = TRUE), 
                                    digits = 3)))
if(K_true > 1){
  print(round(mean(MSDE_Phi[,2], na.rm = TRUE), digits = 3))
}


# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = K_true, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
quantile_mat[1,] = sort(na.omit(MSDE_Phi)[,1])[ceiling(nrow(na.omit(MSDE_Phi)) * 
                                                       c(.05, .50, .95))] 
if(K_true > 1){
  quantile_mat[2,] = sort(na.omit(MSDE_Phi)[,2])[
    ceiling(nrow(na.omit(MSDE_Phi)) * c(.05, .50, .95))]
}

ind_mat = matrix(nrow = K_true, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for (j in 1:3) {
  ind_mat[1,j] = which.min(abs(MSDE_Phi[,1] - quantile_mat[1,j]))
  if(K_true > 1){
    ind_mat[2,j] = which.min(abs(MSDE_Phi[,2] - quantile_mat[2,j]))
  }
}
phi_1_5th = sign(Phi_1_est[ind_mat[1,1],7]) * Phi_1_est[ind_mat[1,1],]
phi_1_50th = sign(Phi_1_est[ind_mat[1,2],7]) * Phi_1_est[ind_mat[1,2],]
phi_1_95th = sign(Phi_1_est[ind_mat[1,3],7]) * Phi_1_est[ind_mat[1,3],]


## plot for phi_1(t)
matplot(tobs, cbind(Phi_true[,1], phi_1_5th, phi_1_50th, phi_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4), 
        mgp = c(2.5,1,0), xaxs = 'i')
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(phi)[1]*(t)), side=2, line=1.9, cex=1.9)


## plot for phi_2(t)
if(K_true > 1){
  phi_2_5th = sign(Phi_2_est[ind_mat[1,1],1]) * Phi_2_est[ind_mat[1,1],]
  phi_2_50th = sign(Phi_2_est[ind_mat[1,2],1]) * Phi_2_est[ind_mat[1,2],]
  phi_2_95th = sign(Phi_2_est[ind_mat[1,3],1]) * Phi_2_est[ind_mat[1,3],]
  matplot(tobs, cbind(Phi_true[,2], phi_2_5th, phi_2_50th, phi_2_95th), 
          type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
          cex.axis = 1.5, col=c("black", "red","black", "blue"), 
          lty = c(1,2,3,4), mgp = c(2.5,1,0), xaxs = 'i')
  title(main = list("", cex = 2.5), adj = 0, line = 1.5)
  mtext(expression(hat(phi)[2]*(t)), side=2, line=1.9, cex=1.9)
}
```



Gamma

```{r}
R_true = 1 # only one unique component is used in simulation scenario 1 to generate Y.

# MSDE

MSDE_Gamma = matrix(data = NA, nrow = length(Gamma_save), ncol = R_true)
Gamma_1_est = matrix(data = NA, nrow = length(Gamma_save), ncol = N)
Gamma_2_est = matrix(data = NA, nrow = length(Gamma_save), ncol = N)
Lambda_est = matrix(data = NA, nrow = length(Lambda_save), ncol = R_true)
for (i in 1:length(Gamma_save)) {
  if(!is.null(Gamma_save[[i]])){ # skipped if no unique components are estimated.
    Gamma_est = Gamma_save[[i]]
    if(ncol(Gamma_est) == 1 & R_true == 1){ # R_true = R_est = 1, correspondence is fixed.
      # Considering possible flipping forms between the estimated and true components, we 
      # take the minimal MSDE value from following difference integration.
      MSDE_Gamma[i,1] = min(
        trapz(tobs, (Gamma_true - Gamma_est)^2) / trapz(tobs, Gamma_true^2), 
        trapz(tobs, (Gamma_true + Gamma_est)^2) / trapz(tobs, Gamma_true^2))
      Gamma_1_est[i,] = Gamma_est
      Lambda_est[i,1] = Lambda_save[[i]]
    }else if(R_true == 1){ # R_est > R_true = 1
      # When the number of unique component estimates returned are greater than the true number, the  
      # one having the lowest MSDE with the true component is referred to as its correspondence.
      for (j in 1:ncol(Gamma_est)) {
        MSDE = min(trapz(tobs, (Gamma_true - Gamma_est[,j])^2) / 
                     trapz(tobs, Gamma_true^2), 
                 trapz(tobs, (Gamma_true + Gamma_est[,j])^2) / 
                   trapz(tobs, Gamma_true^2))
        if(j > 1){
          if(MSDE < MSDE_Gamma[i, 1]){
            MSDE_Gamma[i,1] = MSDE
            Gamma_1_est[i,] = Gamma_est[,j]
            Lambda_est[i,1] = Lambda_save[[i]][j,j]
          }
        }else{
          MSDE_Gamma[i,1] = MSDE
          Gamma_1_est[i,] = Gamma_est[,j]
          Lambda_est[i,1] = Lambda_save[[i]][j,j]
        }
      }
    }else if(ncol(Gamma_est) == 1){ # R_true > R_est = 1
        MSDE_1 = min(trapz(tobs, (Gamma_true[,1] - Gamma_est)^2) / 
                       trapz(tobs, Gamma_true[,1]^2), 
                 trapz(tobs, (Gamma_true[,1] + Gamma_est)^2) / 
                   trapz(tobs, Gamma_true[,1]^2))
        MSDE_2 = min(trapz(tobs, (Gamma_true[,2] - Gamma_est)^2) / 
                       trapz(tobs, Gamma_true[,2]^2), 
                 trapz(tobs, (Gamma_true[,2] + Gamma_est)^2) / 
                   trapz(tobs, Gamma_true[,2]^2))
        if(MSDE_1 < MSDE_2){
          MSDE_Gamma[i,1] = MSDE_1
          Gamma_1_est[i,] = Gamma_est
          Lambda_est[i,1] = Lambda_save[[i]]
          MSDE_Gamma[i,2] = NA
          Lambda_est[i,2] = NA
        }else{
          MSDE_Gamma[i,2] = MSDE_2
          Gamma_2_est[i,] = Gamma_est
          Lambda_est[i,2] = Lambda_save[[i]]
          MSDE_Gamma[i,1] = NA
          Lambda_est[i,1] = NA
        }
    }else{ # R_true > 1, R_est > 1.
      ind_gamma_1 = 0 # used to save the index of estimated component corresponding to gamma_1(t)
      MSDE_array = c()
      for (j in 1:ncol(Gamma_est)) {
        MSDE_array = append(MSDE_array, min(
          trapz(tobs, (Gamma_true[,1] - Gamma_est[,j])^2) / 
            trapz(tobs, Gamma_true[,1]^2),
          trapz(tobs, (Gamma_true[,1] + Gamma_est[,j])^2) / 
            trapz(tobs, Gamma_true[,1]^2)))
        ind_gamma_1 = which.min(MSDE_array)
      }
      
      ind_gamma_2 = 0 # used to save the index of estimated component corresponding to gamma_2(t)
      MSDE_array = c()
      for (j in 1:ncol(Gamma_est)) {
        MSDE_array = append(MSDE_array, min(
          trapz(tobs, (Gamma_true[,2] - Gamma_est[,j])^2) / 
            trapz(tobs, Gamma_true[,2]^2), 
          trapz(tobs, (Gamma_true[,2] + Gamma_est[,j])^2) / 
            trapz(tobs, Gamma_true[,2]^2)))
        ind_gamma_2 = which.min(MSDE_array)
      }
      
      if(ind_gamma_1 != ind_gamma_2){
        MSDE_Gamma[i, 1] = min(
          trapz(tobs, (Gamma_true[,1] - Gamma_est[,ind_gamma_1])^2) / 
            trapz(tobs, Gamma_true[,1]^2),
          trapz(tobs, (Gamma_true[,1] + Gamma_est[,ind_gamma_1])^2) / 
            trapz(tobs, Gamma_true[,1]^2))
        Gamma_1_est[i,] = Gamma_est[,ind_gamma_1]
        Lambda_est[i,1] = Lambda_save[[i]][ind_gamma_1,ind_gamma_1]
        
        MSDE_Gamma[i, 2] = min(
          trapz(tobs, (Gamma_true[,2] - Gamma_est[,ind_gamma_2])^2) / 
            trapz(tobs, Gamma_true[,2]^2), 
          trapz(tobs, (Gamma_true[,2] + Gamma_est[,ind_gamma_2])^2) / 
            trapz(tobs, Gamma_true[,2]^2))
        Gamma_2_est[i,] = Gamma_est[,ind_gamma_2]
        Lambda_est[i,2] = Lambda_save[[i]][ind_gamma_2,ind_gamma_2]
      }else{
        print(paste0("In ", i, "th Monte Carlo run, the two true unique ", 
                     "components are corresponding to the same estimate."))
      }
    }
  }
}

print(paste0("The mean MSDE for the unique components {", symbol("gamma"), 
             "(t)} ", "are:", round(mean(MSDE_Gamma[,1], na.rm = TRUE), 
                                    digits = 3)))
if(R_true > 1){
  print(round(mean(MSDE_Gamma[,2], na.rm = TRUE), digits = 3))
}


# Plots

## Choose the estimate with (5th, 50th, 95th) MSDE with the true latent component
quantile_mat = matrix(nrow = R_true, ncol = 3) # a matrix saves the (5th, 50th, 95th) MSDE values.
quantile_mat[1,] = sort(na.omit(MSDE_Gamma)[,1])[ceiling(nrow(na.omit(
  MSDE_Gamma)) * c(.05, .50, .95))] 
if(R_true > 1){
  quantile_mat[2,] = sort(na.omit(MSDE_Gamma)[,2])[ceiling(nrow(na.omit(
    MSDE_Gamma)) * c(.05, .50, .95))]
}

ind_mat = matrix(nrow = R_true, ncol = 3) # a matrix contains the indices corresponding to the selected MSDE
for (j in 1:3) {
  ind_mat[1,j] = which.min(abs(MSDE_Gamma[,1] - quantile_mat[1,j]))
  if(R_true > 1){
    ind_mat[2,j] = which.min(abs(MSDE_Gamma[,2] - quantile_mat[2,j]))
  }
}
gamma_1_5th = sign(Gamma_1_est[ind_mat[1,1],4]) * Gamma_1_est[ind_mat[1,1],]
gamma_1_50th = sign(Gamma_1_est[ind_mat[1,2],4]) * Gamma_1_est[ind_mat[1,2],]
gamma_1_95th = sign(Gamma_1_est[ind_mat[1,3],4]) * Gamma_1_est[ind_mat[1,3],]

## plot for gamma_1(t)
matplot(tobs, cbind(Gamma_true[,1], gamma_1_5th, gamma_1_50th, gamma_1_95th), 
        type = "l", lwd = 3, xlab = "Time", ylab = "", cex.lab = 1.9, 
        cex.axis = 1.5, col=c("black", "red","black", "blue"), lty = c(1,2,3,4),
        mgp = c(2.5,1,0), xaxs = 'i')
title(main = list("", cex = 2.5), adj = 0, line = 1.5)
mtext(expression(hat(gamma)[1]*(t)), side=2, line=1.9, cex=1.9)


## plot for gamma_2(t)
if(R_true > 1){
  gamma_2_5th = sign(Gamma_2_est[ind_mat[1,1],4]) * Gamma_2_est[ind_mat[1,1],]
  gamma_2_50th = sign(Gamma_2_est[ind_mat[1,2],4]) * Gamma_2_est[ind_mat[1,2],]
  gamma_2_95th = sign(Gamma_2_est[ind_mat[1,3],4]) * Gamma_2_est[ind_mat[1,3],]
  matplot(tobs, cbind(Gamma_true[,2], gamma_2_5th, gamma_2_50th, gamma_2_95th), 
          type = "l", lwd = 3, xlab = "Time", ylab = "",  cex.lab = 1.9, 
          cex.axis = 1.5, col=c("black", "red","black", "blue"), 
          lty = c(1,2,3,4), mgp = c(2.5,1,0), xaxs = 'i')
  title(main = list("", cex = 2.5), adj = 0, line = 1.5)
  mtext(expression(hat(gamma)[2]*(t)), side=2, line=1.9, cex=1.9)
}
```


Calculate MSE of score variances

```{r}
scoreVar_MSE = matrix(data = NA, nrow = length(Omega_x_save), ncol = 4)
for (i in 1:length(Omega_x_save)) {
    scoreVar_MSE[i,1] = (diag(Omega_x) - Omega_x_est[i,])^2 / (diag(Omega_x))^2
    scoreVar_MSE[i,2] = (diag(Omega_y) - Omega_y_est[i,])^2 / (diag(Omega_y))^2
    scoreVar_MSE[i,3] = (diag(Theta) - Theta_est[i,])^2 / (diag(Theta))^2
    scoreVar_MSE[i,4] = (diag(Lambda) - Lambda_est[i,])^2 / (diag(Lambda))^2
}

print("Mean MSE for the following score variances are:")
print(paste(symbol("omega") %p% supsc("x"), ":", 
            round(mean(scoreVar_MSE[,1], na.rm = TRUE), digits = 3)))
print(paste(symbol("omega") %p% supsc("y"), ":", 
            round(mean(scoreVar_MSE[,2], na.rm = TRUE),digits = 3)))
print(paste(symbol("theta"), ":", 
            round(mean(scoreVar_MSE[,1], na.rm = TRUE),digits = 3)))
print(paste(symbol("lambda"), ":", 
            round(mean(scoreVar_MSE[,1], na.rm = TRUE),digits = 3)))
```


